---
title: "p8105_hw2_qz2266"
author: "Qing Zhou"
date: "2022-10-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r libraries}
library(tidyverse)
library(readxl)
```

# Problem 2

## Read and clean the Mr. Trash Wheel sheet.

```{r}
mr_trash_wheel = 
  read_excel("data/Trash Wheel Collection data-new.xlsx", 
  sheet = "Mr. Trash Wheel", 
  range =  cell_cols("A:N"), skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls, digits = 0)),year = as.integer(year))
  
```

## Read and clean the Professor Trash Wheel sheet.

```{r}
prof_trash_wheel = 
  read_excel("data/Trash Wheel Collection data-new.xlsx", 
  sheet = "Professor Trash Wheel", 
  range =  cell_cols("A:M"), skip = 1) %>% 
  
  janitor::clean_names() %>% 
  drop_na(dumpster) 
```

## Add a new variable to keep track of which Trash Wheel is which.

```{r}
mr_trash_wheel_new = mutate(mr_trash_wheel, trash_wheel_type = "Mr.") %>%
                     relocate(trash_wheel_type, .before = dumpster) 
                     
prof_trash_wheel_new = mutate(prof_trash_wheel, trash_wheel_type = "Prof.") %>% 
                       relocate(trash_wheel_type, .before = dumpster) 
```                       

## Combine two datasets

```{r}
trash_wheel = 
  bind_rows(mr_trash_wheel_new, prof_trash_wheel_new)
```

## Data discription

The original mr_trash_wheel data set has `r ncol(mr_trash_wheel)` variables, including `r colnames(mr_trash_wheel)`, and `r nrow(mr_trash_wheel)` observations. 
The original prof_trash_wheel data set has `r ncol(prof_trash_wheel)` variables, including `r colnames(prof_trash_wheel)`, and `r nrow(prof_trash_wheel)` observations. 
As for the combined data set, there are `r ncol(trash_wheel)` variables, and `r nrow(trash_wheel)` observations. Key variables are measurements of trash collected, including weight,volume and trash type, such as plastic bottles, cigarette butts, grocery bags -- in addition to the date of the collection.


The total weight of trash collected by Professor Trash Wheel is `r round(sum(pull(prof_trash_wheel, weight_tons), na.rm = TRUE))` tons. Similarly, if we use the combined data set, it would be `r round(sum(pull(filter(trash_wheel, trash_wheel_type == 'Prof.'),  weight_tons), na.rm = TRUE))` tons.


The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r round(sum(pull(filter(mr_trash_wheel, year == 2020), sports_balls), na.rm = TRUE))`. If we use the combined data set, the number is the same. It is `r round(sum(pull(filter(trash_wheel, trash_wheel_type == 'Mr.' & year == 2020),  sports_balls), na.rm = TRUE))`.


# Problem 3

## Clean the data in pols-month.csv.

```{r read_pols_month}
pols = read_csv("data/fivethirtyeight_datasets/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  
  separate(col = mon, into = c("year", "month", "day")) %>%
  mutate(
    year = as.integer(year), 
    day = as.integer(day),
    month = month.name[as.integer(month)]) %>%
  
  mutate(president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>%
  select(-prez_gop, -prez_dem, -day)
```

## Clean the data in snp.csv.

```{r read_snp}

library(lubridate)

snp = read_csv("data/fivethirtyeight_datasets/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  
  mutate(date = lubridate::parse_date_time2(date, orders = "mdy", cutoff_2000 = 23)) %>%
  separate(col = date, into = c("year", "month", "day")) %>%
  mutate(
    year = as.integer(year), 
    day = as.integer(day)) %>%
  arrange(year, month) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  
  select(year, month, day, close)
```

## Clean the data in unemployment.csv.

 
```{r read_unemployment}
unemp =
  read_csv("data/fivethirtyeight_datasets/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp_rate"
  ) %>%
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  janitor::clean_names() %>%
  mutate(year = as.integer(year))
```

## Merge the data sets

```{r merge}
snp_pols_merge = 
  left_join(pols, snp, by = c("year","month")) %>%
  relocate(year, month, day)

merge_538 = 
  left_join(snp_pols_merge, unemp, by = c("year","month"))
  
```

## Dpescription of the dataset

The pols-month dataset describes the political landscape of the US, such as the number of politicians who are democratic or republican at the presidential, governor, and senate level on the associated date. It has  `r ncol(pols)` variables, including `r colnames(pols)`, and `r nrow(pols)` observations, from `r min(pull(pols, year))` to `r max(pull(pols, year))`.


The snp dataset describes the closing vales of the S&P stock index on the associated date. It has  `r ncol(snp)` variables, including `r colnames(snp)`, and `r nrow(snp)` observations, from `r min(pull(snp, year))` to `r max(pull(snp, year))`.


The unemployment data set introduces the percentage of unemployment rate in each month of the associated year. It has  `r ncol(unemp)` variables, including `r colnames(unemp)`, and `r nrow(unemp)` observations, from `r min(pull(unemp, year))` to `r max(pull(unemp, year))`.


The final dataset merge_538 combines the previous 3 datasets together, by year and month, creating a large dataset that contains information on the political landscape, stock performance, and unemployment rate from `r min(pull(merge_538, year))` to `r max(pull(merge_538, year))`. This final dataset dimensions are  `r nrow(merge_538)` rows and `r ncol(merge_538)` columns, including `r colnames(merge_538)`.


One key variable in this dataset is the **president** variable. It indicates whether the president on the associated date was republican or democratic. This variable could be used to evaluate the performance of the president, when considering other variables.


